---
title : ISTA 320 Final
author: Robert Lopez  
email : rlopez4747@email.arizona.edu
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
# loading data and libraries
library(tidyverse)
library(ggplot2)
library(reshape2)
library(dplyr)
library(knitr)
performance_df <- read_csv('StudentsPerformance.csv')

# remove race/ethnic column, we are not gonna use it. 
performance_df <- performance_df %>%
  select(-`race/ethnicity`)


# get rid of spaces and apostrophes 
performance_df$`parental level of education` <- performance_df$`parental level of education` %>%
  str_replace('\\s', '_') %>%
  str_remove('[\']')
```

Question One: Which exam gets the highest scores?
```{r, include = FALSE}
performance_melted <- performance_df %>%
  select(-`parental level of education`, -lunch, -`test preparation course`, -gender)

performance_melted <- melt(as.data.frame(performance_melted))
```

The following histograms show the scores from 90 percent to 100 percent. The reading scores have the most 100 percents. This means that this exam gets the higher scores overall. It must be easier. 
```{r,  include=TRUE}
ggplot(performance_melted %>% filter(value > 90),
       aes( x = value,
            fill = variable)) +
  geom_bar() +
  facet_wrap(~variable) +
  labs(title = 'Total Exam Scores From 90-100%',
       x = 'Percent',
       y = 'Total')
```

Question two: How many students completed the test preparation?
The following pie chart shows how many students completed the test preparation course.

```{r}
test_prep <- performance_df %>%
  select(`test preparation course`)

test_prep$`test preparation course` <- ifelse(test_prep$`test preparation course` == 'none', 0, 1)

no <- sum(test_prep$`test preparation course` == 0)
yes <- sum(test_prep$`test preparation course` == 1)
total <- c(no, yes)
lbls <- c('not-taken', 'taken')
piepercent <- round(100*total/sum(total), 1)
pie(total, labels = piepercent, main = 'Students who completed Test Prep' ,col = c('#E1BE6A', '#40B0A6')) 

legend("topright", c("Not Taken","Taken"), cex = 0.8,
   fill = c('#E1BE6A', '#40B0A6'))
```

Question Three: Did the students who completed the test prep score higher than the ones who did not?
